---
title: "Covid Cases per state by political affiliation"
author: "Ayush Arora"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(maps)
library(ggmap)
```



#### Importing political affiliation by state data from CSV
```{r}
df = read.csv("https://raw.githubusercontent.com/Araisedtotwo/Final-Project/main/raw_data.csv",skip = 2, header = T)
```

#### Cleaning the data - Removing extra columns
```{r}
df$State.Senate.Majority.Political.Affiliation <- NULL
df$State.House.Majority.Political.Affiliation <- NULL
df$State.Attorney.General.Political.Affiliation  <- NULL
df$State.Insurance.Commissioner.Political.Affiliation  <- NULL
df$Footnotes   <- NULL
head(df,5)
```

#### Renaming column "State.Code" to "state" and

#### "Governor.Political.Affiliation" to  "affiliation" for join
```{r}
df2 <- df %>% 
  rename(
    state = State.Code,
    affiliation = Governor.Political.Affiliation
    )
head(df2,5)
```
#### View Invalid rows at the bottom of CSV
```{r}
tail(df2,4)
```

#### Cleaning the data - Removing invalid rows from the bottom (not part of the analysis)
```{r}
cleaned <- head(df2,51)
head(cleaned,5)
```
### Data Source 2 - API
#### Extracting COVID data from API for merging with political affiliation dataset
```{r warning=FALSE}
cases <- fromJSON("https://data.cdc.gov/resource/9mfq-cb36.json?submission_date=2020-10-31", flatten = TRUE)
head(cases,5)
```
#### Formatting Date
```{r warning=FALSE}
cases[[1]] <- substr(cases$submission_date, 0, 10)
head(cases,5)
```

#### Joining/Merging the dataframes
```{r}
final_df <- merge(cleaned, cases)
head(final_df, 5)
```

#### Converting total_cases and total_deaths to numeric
```{r}
final_df <- transform(final_df, tot_cases = as.numeric(tot_cases))
final_df <- transform(final_df, tot_death = as.numeric(tot_death))
head(final_df, 5)
```
#### Extracting data for the Republican States
```{r}
republican <- final_df[final_df$affiliation == 'Republican',]
head(republican,3)
```

#### Extracting data for the Democrat States
```{r}
democrat <- final_df[final_df$affiliation == 'Democrat',]
head(democrat,3)
```
#### Total cases for republican states
```{r}
republican$tot_cases
length(republican$tot_cases)
```

#### Total cases for Democrat states
```{r}
democrat$tot_cases
length(democrat$tot_cases)
```

#### Summary for Democratic States
```{r}
summary(democrat$tot_cases)
```
#### Summary for Republican States
```{r}
summary(republican$tot_cases)
```
#### Box Plot : Covid Cases Distribution by states' political party affiliation
##### Republicans (1) and Decomcrats (2)

```{r}
dat_ggplot <- data.frame(
  value = c(republican$tot_cases,democrat$tot_cases),
  Political_Affiliation = c(rep("Republican States",26), rep("Democrat States", 25))
)

ggplot(dat_ggplot) +
  aes(x = Political_Affiliation, y = value) +
  geom_boxplot() +
  theme_minimal()
```

#### T-test: Check whether average total cases in Democrat vs Republican affiliated states are significantly different
```{r}
t.test(republican$tot_cases, y = democrat$tot_cases,alternative = c("two.sided"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```


#### Conclusion 1

- As can be seen from the box plot and the descriptive statistics, the median number of Covid Cases are nearly the same for the states affiliated to either party.
- Since the p-value of 0.729 is greater than the 5% level of significance from the T-test, the average number of cases in Republican states are not significantly different from the average number of cases in the Democrat states.

---

#### Analysis 2: Finding difference in average death proportion for states of different political party affiliation

---

#### Computing death proportion (percentage)
```{r}

final_df$death_percent <-  final_df$tot_death / final_df$tot_cases * 100

final_df %>%
  select(Location, affiliation, tot_death, tot_cases,death_percent)

```

#### Segregating affiliations for T-test
```{r}
republican2 <- final_df[final_df$affiliation == 'Republican',]
democrat2 <- final_df[final_df$affiliation == 'Democrat',]
summary(republican2$death_percent)
summary(democrat2$death_percent)
```
#### T-test for death proportion
```{r}
t.test(republican2$death_percent, y = democrat2$death_percent,alternative = c("two.sided"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

### Death Percantage Visualizations
```{r}
Republican_Death_Percentage <- republican2$death_percent
hist(Republican_Death_Percentage)
```
```{r}

Democrat_Death_Percentage <- democrat2$death_percent
hist(Democrat_Death_Percentage)
```

```{r}

usa <- map_data("state")
```
#### Additional Feature : Plotting Death proportions using ggmap
```{r}
final_df2<-final_df[!(final_df$Location=="Alaska" | final_df$Location=="Hawaii"),]

ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_point(data = final_df2, size=final_df2$death_percent, mapping = aes(x = longitude, y = latitude), color = ifelse(final_df2$affiliation =="Republican",'red','blue'))
```
--- 

#### Conclusion 2

---

- Since the p-value of 0.0753 is marginally greater than the 5% level of significance from the T-test, the average death percentage in Republican states are not significantly different from the average number of cases in the Democrat states.

- But, from the histogram (highest length in the last bucket) and map, it can be seen that states affiliated with Democrats have higher death percentage than those affiliated with Republicans.